---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/README-", 
  fig.width = 8.2, 
  fig.height = 5.5
  )

```

 
```{r global_options, echo = FALSE}
knitr::opts_chunk$set(fig.align = 'center', echo=FALSE, warning=FALSE, message=FALSE, results = "hide", dev = 'png')
```

eeg-complex
-------
This package contains code that reproduces figures and analyses used in my MA thesis, _Improvement of epsilon-complexity estimation and an application to seizure prediction_.

For an overview of the research and individual analyses see the [web pages](https://nateaff.github.io/eeg-complex/). 

The [ecomplex](https://github.com/nateaff/ecomplex) package contains the `R` version of the procedure for computing the epsilon-complexity of a single variable function or time series. The repositories for the three other packages used in this thesis, [tssims](https://github.com/nateaff/tssims), [tsfeats](https://github.com/nateaff/tsfeats) and the imaginatively named [tssegment](https://github.com/nateaff/tssegment), are also on github. The first two packages unify access to a set of simulations and features, respectively. The `tssegment` package provides an interface to the segmentation and classification methods used for predicting seizures. 

The current draft of the complete thesis is [here.](https://github.com/nateaff/eeg-complex/blob/master/docs/thesis/thesis.pdf) (This is a draft and has not yet been submitted.)

## Files 

This is an `R` package with additional files that follow the structure of a [workflowr](https://github.com/jdblischak/workflowr) project. See the [workflowr documentation](https://jdblischak.github.io/workflowr/index.html) for more information. The [workflowr readme](https://github.com/jdblischak/workflowr) also as a nice overview of similar project management packages.

`R` : Contains most standalone functions. 

`code`: Code associated with individual analyses. Code is organized by prefix into 5 sections:

  * `sim`,`ecomplex`: Simulations and testing of the `ecomplex` procedure.
  * `holder`: Test of the properties of the epsilon-complexity coefficients and their relation to Holder continuity and fractal dimension.
  * `eeg-prep`: Data cleaning and feature extraction from the raw eeg data.
  * `eeg`: Prediction and analysis of the derived eeg features. 

`analysis`: These are Rmarkdown files used to build the accompanying web pages. The analyses provide samples of the results from each section. The data pre

`docs`: The web-pages and related documents such as slide presentations and the thesis tex and pdf files. 

`output`: Output of intermediate analyses 




