---
title: "Properties of the Complexity Coefficients"
author: "Nathanael Aff"
date: 2017--07-10
output: 
  html_document:
    toc: true
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```


<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

```{r, holder-weir0}
  devtools::load_all()
  knitr::read_chunk("../code/holder-weierstrass-notrandom.R")
```
### Introduction

Darkhovsky and Piryatinska have conjectured that for a constant generating process the complexity coefficients time series generated by that process are constant. More concretely, we tested if, for a constant Hölder class of functions, the complexity coefficients are also constant. 

A Hölder-$\alpha$ class of functions is a set of functions with a bounded rate of change indexed by the parameter $\alpha$. Specifically, alpha Hölder continuous function $f(x)$ is one that satisfies
\[
    | f(x + h) - f(x)| < C|h|^{\alpha},
\]
for positive constants $C$, $\alpha$.

Three of the simulations used in our previous tests have parameters that determine the Hölder-$\alpha$ class of the function or time series generated by the process -- fractional Brownian Motion (fBm), the random Weierstrass function and the Cauchy process. Along with the standard Weierstrass function, the complexity coefficients were estimated for time series generated by a range of Hölder class determining parameters. For all of the tested functions or random processes, the Hölder class and fractal dimension were determined by the same parameter. 

For all process except for the random Weierstrass function, the mean of the complexity coefficient $B$ changed linearly with the Hölder class of the time series and with the time series' estimated fractal dimension. Here we present the results from the Weierstrass function and the Cauchy process.

### Weierstrass Function

The Weierstrass function is well-known function that is continuous but nowhere differentiable. The function can be written:
$$
 W_{\alpha}(x) = \sum_{n=0}^{\infty}b^{-n\alpha}\cos(b^n \pi x).
$$
In this form, the single parameter $\alpha$ is the Hölder class for the function and $2-\alpha$ is the functions fractal dimension[^1]. A plot of the Weierstrass function for several parameters is shown below.


```{r, holder-weir1, fig.cap = "Plot of Weierstrass function with alpha parameters (0.2, 0.42, 0.8)"}

<<weierstrass-notrandom-plot>>
```

The complexity coefficients and fractal dimension were estimated for 20 values of $\alpha$. The results plotted below show that the all both complexity coefficients and fractal dimension change linearly with the $\alpha$ parameter.

```{r holder-weir2, fig.height = 3, fig.cap = "Complexity coefficients and fractal dimension plotted against the Weierstrass parameter alpha."}
  
  <<holder-coefficients>> 

```

```{r, cauchy0}
  knitr::read_chunk("../code/holder-cauchy-plots.R")
```

### Cauchy process 

Fractional Brownian motion has a single parameter that determines both the Hölder class/fractal dimension of its sample paths and the Hurst parameter. The Hurst parameter is a measure of the long range dependence of a random process. On the other hand, the Cauchy process has two parameters which affect the fractal dimension and Hurst parameter of sample paths of the process in isolation. The Cauchy process is defined in terms of its correlation function 
\[
   c(h) = (1 + |h|^{\alpha})^{-\beta/\alpha}, \hspace{1em} h \in \mathbb{R}.
\]

For $\alpha \in (0, 2]$ its fractal dimension is
\[
 D = 2 - \frac{\alpha}{2}. 
\] 

And for $\beta \in (0,1)$ its Hurst parameter is 
\[
    H = 1 - \frac{\beta}{2}.
\]

The plot below shows the Cauchy process on a 3 $\times$ 3 grid of parameters. Fractal dimension (and the Hölder class) are constant on the columns.

```{r cauchy1, fig.height = 7, fig.width = 9, fig.cap = "Plot of Cauchy process for a grid of parameters."}
  <<cauchy>>
```

30 samples were generated for each parameter combination on a 10 $\times$ 10 grid. The plots below show the estimates for the complexity coefficients and fractal dimension plotted against the $\alpha$ parameter and $\beta$ parameter. The plots show the complexity coefficient $B$ and the fractal dimension estimator change linearly with $\alpha$ but are largely unaffected by the changes in the Hurst parameter.

```{r, cauchy2}
  knitr::read_chunk("../code/holder-cauchy.R")
```

```{r cauchy3, fig.height = 3, fig.cap = "Plot of complexity coefficients and fractal dimension of the Cauchy processes generated from a grid of parameters."}
  <<cauchy-coefficients>>
```

### Discussion 

Similar results held for fBm but not for the random Weierstrass function. 
For each process tested in this section, the complexity coefficient $B$ behaved like the variogram-based fractal dimension estimator.

It's likely that the complexity coefficient $B$ is an estimator of fractal dimension. One approach to proving this is would be to restrict the proof, for example, to stationary Gaussian processes.


[^1]: This was recently proved by [Shen](https://arxiv.org/abs/1505.03986).

<!-- Add your analysis here -->

### Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
