---
title: "Epsilon-complexity Procedure"
author: "Nathanael Aff"
date: 2017-07-15
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

The practical estimation of $\varepsilon$-complexity uses some set of approximation methods to successively downsample and approximation a function. At each $h$th step, only some proportion of the initial samples, $S_h$, are used finding the minimum approximation error,  $\varepsilon_h$. A least squares fit 
\[
  \log(S) \approx A + B\log(\varepsilon)
\]
determines the two parameters $A$ and $B$ which characterize the $\varepsilon-$complexity of the function. 

<hr>
### $\varepsilon$-complexity algorithm

<hr>
**Input**: $X$ a regularly sampled time series of length $N$. <br>
**Input**: $\mathcal{F}$ a set of approximation methods $f$. <br>
**Input**: $\mathcal{H}$ the set of spacings $h$. <br>
**Output**: The complexity coefficients $A,B$. <br>
<hr>

**foreach** $h$ in $\mathcal{H}$ **do** 
<br>
<p style="margin-left: 20px">**foreach** $f$ in $\mathcal{F}$ **do** 
</p>
<p style="margin-left: 40px">Compute the approximation error <br> 
        $\varepsilon_{h,f} \leftarrow \frac{1}{N}
         \left|f_h - X_{h} \right|^2$ 
        <br>
</p>
<p style="margin-left: 20px">**end** <br>
     Find minimium error over all $f$  <br>
     epsilons$_h$ $\leftarrow \min \varepsilon_{h,f}$. 
</p>
**end** <br>
Fit a least squares linear model <br>
$A,B$ $\leftarrow$ lm$\left(\{ S_h \} \sim \{ \varepsilon_h \} \right)$.
<br>

<hr>
