---
title: "Properties of the Complexity Coefficients"
author: "Nathanael Aff"
date: 2017--07-10
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```


```{r, holder-weir0}
  devtools::load_all()
  knitr::read_chunk("../code/holder-weierstrass-notrandom.R")
```

Darkhovsky and Piryatinska have conjectured that for a constant generating process the complexity coefficients are constant. More concretely, we tested if, for a constant Hölder class of functions the complexity coefficients are also constant. 

The Weierstrass function is well-known function that is continuous but nowhere differentiable. The function can be written:
$$
 W_{\alpha}(x) = \sum_{n=0}^{\infty}b^{-n\alpha}\cos(b^n \pi x).
$$
In this form, the single parameter $\alpha$ is the Hölder class for the function and $2-\alpha$ is the functions fractal dimension[^1]. A plot of the Weierstrass function for several parameters is shown below.


```{r, holder-weir1, fig.cap = "Plot of Weierstrass function with alpha parameters (0.2, 0.42, 0.8)"}

<<weierstrass-notrandom-plot>>
```

The complexity coefficients and fractal dimension was estimated for 20 values of $\alpha$. The results plotted below show that the all both complexity coefficients and fractal dimension change linearly with the $\alpha$ parameter.

```{r holder-weir2, fig.height = 3, fig.cap = "Complexity coefficients and fractal dimension plotted against the Weierstrass parameter alpha."}
  
  <<holder-coefficients>>

```

```{r, cauchy0}
  knitr::read_chunk("../code/holder-cauchy-plots.R")
```

Several of the random processes plotted [here](sim-tests.html) have parameters that determine both the Hölder class of samples paths generated by the process. For the random-Weierstrass function, fractional Brownian motion(fBm), and the Cauchy process, a single parameter determines both the Hölder class and the fractal dimension of the process. 

Fractional Brownian motion has a single parameter that determines both the Hölder class/fractal dimension of its sample paths and the Hurst parameter. The Hurst parameter is a measure of the the long range dependence of a random process. 

On the other hand, the Cauchy process has two parameters which isolate the fractal dimension and Hurst parameter. The Cauchy process is defined in terms of its correlation function 
\[
   c(h) = (1 + |h|^{\alpha})^{-\beta/\alpha}, \hspace{1em} h \in \mathbb{R}.
\]

For $\alpha \in (0, 2]$ its fractal dimension is
\[
 D = 2 - \frac{\alpha}{2}. 
\] 

And for $\beta \in (0,1)$ its Hurst parameter is 
\[
    H = 1 - \frac{\beta}{2}.
\]

The plot below shows the Cauchy process on a 3 $\times$ 3 grid of parameters. Fractal dimension (and the Hölder class) are constant on the columns.

```{r cauchy1, fig.height = 7, fig.width = 9, fig.cap = "Plot of Cauchy process for a grid of parameters."}
  <<cauchy>>
```

30 samples were generated for each parameter combination on a 10 $\times$ 10 grid. The plots below show the estimates for the complexity coefficients and fractal dimension plotted first gainst the $\alpha$ parameter and secondly against the $\beta$ parameter. The plots show the complexity coefficient $B$ and the fractal dimension estimator change linearly with $\alpha$ but are largely unaffected by the changes in the Hurst parameter.

```{r, cauchy2}
  knitr::read_chunk("../code/holder-cauchy.R")
```


```{r cauchy3, fig.height = 3, fig.cap = "Plot of complexity coefficients and fractal dimension on a grid Cauchy process sampled from a grid of parameters."}
  <<cauchy-coefficients>>
```

Similar results held for fBM but not for the random Weierstrass function. 
In all cases, the complexity coefficient behaved similar to the fractal dimension estimator based on the empirical variogram of the function. 

It's likely that the complexity coefficient $B$ is an estimator of fractal dimension. One approach to proving this is would be to restrict the proof to, for example, stationary Gaussian processes.


[^1]: This was recently proved by [Shen](https://arxiv.org/abs/1505.03986).

<!-- Add your analysis here -->

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
