---
title: "Project overview"
output:
  html_document:
    toc: false
    code_folding: hide
---


<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("../analysis/chunks.R")

```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}

```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}

```
<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->

1. [Simulation and Approximation Methods](simulations.html) 
2. [Properties of the Complexity Coefficients](complexity-coefficients.html)
3. [Seizure Prediction](seizure-prediction.html)
4. [Computing the Complexity Coefficients](ecomplexity-procedure.html)

The $\varepsilon$-complexity of a continuous function is, roughly speaking, the amount of information needed to reconstruct that function within an absolute error $\varepsilon$. The theory and procedure for estimating $\varepsilon$-complexity was introduced by Darkhovsky and Piryatinska[^1]. In previous work, the complexity coefficients have been shown to be useful features for the segmentation and classification of time series[^2]. 

The purpose of these analyses was three-fold: 

1. To improve the estimation of the complexity coefficients by enlarging the set of approximation methods used in the estimation procedure.

2. To explore the relation between the complexity coefficients, the Hölder condition and fractal dimension. 

3. An application of the complexity coefficients to the segmentation and classification of EEG.

4. Along with an overview of the results of each of these analyses we outline the procedure for estimating the complexity coefficients.


```{r, segment-plot1}
library(devtools)
load_all()
knitr::read_chunk("../code/eeg-segment-plot.R")
```


```{r segment-plot2, fig.align = "center", fig.width = 5, fig.height = 3.6, fig.cap = "Segmentation of spectral freatures based complexity coefficient change points"}
  par(mar = c(0, 0, 2, 0))
  <<eeg-segment-create>> 
  <<eeg-segment-plot2>>
  
```
<br>

### Credits 

These poject pages were developed with the help of the [workflowr](https://jdblischak.github.io/workflowr/) package.


[^1]: [Epsilon-complexity of continuous functions](https://arxiv.org/abs/1303.1777)
[^2]: [Binary classification of multi-channel EEG records based on the ε-complexity of continuous vector functions](https://arxiv.org/abs/1610.01633)