---
title: "Set up"
author: "Nathanael Aff"
date: 2017-07-24
output: 
  html_document:
    code_folding: show
    toc: true
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r , knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

The associated [repository](https://github.com/nateaff/eeg-complex) is an `R` package and installation will import a number of additional packages. The analysis also requires three additional packages used as time series analysis helpers. You will also need the `workflowr` and `devtools` packages.

```{r eval=FALSE}
  install.packages("devtools")
  devtools::install_github("jdblischak/workflowr")
  # The project repository
  devtools::install_github("nateaff/eeg-complex")
  # The time series analysis packages
  devtools::install_github(c("nateaff/tssims", "nateaff/tsfeats", "nateaff/tssegment"))

```

Set the working directory to the project folder and load the package. 

```{r eval=FALSE}
# Load project package
devtools::load_all()
```

The project pages can be built and viewed with `workflowr`. The pages will only be re-built if there are changes to the file. The individual analyses have a flag that indicates whether cached intermedidate data should be used in building the analysis. To build files from scratch set the `from_cache` flag to `FALSE`.

```{r eval=FALSE}
build_from_cache(FALSE)
```

This will only work for the files without the `eeg-` prefix. The `eeg` data files are not included in the repo so those files can't be built from scratch, sorry :(.

```{r eval=FALSE}
# This only builds the project analyses
workflowr::wflow_build()
```
The full thesis is the `docs/thesis` and can be built from the `thesis.tex` file.

